---
title: "PA1_template.Rmd"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



### Loading the data:
```{r}
if(!file.exists("activity.csv")) {
        tempfile <- tempfile()
        download.file("http://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip",destfile = tempfile)
        unzip(tempfile)
        unlink(tempfile)
}
```


##### Libraries needed to run the script:
```{r}
library(datasets)

library(dplyr)

library(data.table)

library(ggplot2)

library(xtable)

library(ggplot2)

library(lubridate)

```

### Reading the data:
```{r}
activity_data <- read.csv("activity.csv")
head(activity_data)
```
 
##### Activity names:
```{r}
activity_names <- names(activity_data)
print(activity_names)
```

##### Data dimensions:
```{r}

dim(activity_data)
print(dim(activity_data))
```

##### Summary:
```{r}
summary(activity_data$steps)
```

##### Structure of data:
```{r}
str(activity_data)
```


### Total steps taken each day:
```{r}
steps_day <- aggregate(steps ~ date, data = activity_data, FUN = sum, na.rm = TRUE)
hist(steps_day$steps, xlab = "Steps per Day", main = "Total Steps per Day", col = "blue")

```

### Mean and median number of steps taken each day:
```{r}
mean_steps <- mean(steps_day$steps)
mean_steps <- format(mean_steps,digits=1)

median_steps <- median(steps_day$steps)
median_steps <- format(median_steps,digits=1)
```

```{r}
print(mean_steps)
print(median_steps)
```

### Time series plot of the average number of steps taken:
```{r}
activity_mean <- aggregate(steps ~ interval, data = activity_data, FUN = mean, na.rm = TRUE)
plot(activity_mean$interval, activity_mean$steps, type = "l", col = "red", xlab = "Intervals", ylab = "steps per interval", main = "Averaged Steps per Interval (NA removed)")
```

### The time interval containing the maximum number of steps:

```{r}
max_steps <-max(activity_mean$steps)
print(max_steps)
```

##### Maximum steps during given interval with two significan figures after decimal point:
```{r}
max_interval <- activity_mean$interval[which(activity_mean$steps == max_steps)]
max_steps <- round(max_steps, digits = 2)
print(max_steps)
```

### Missing data

##### Sum of NAs in activity:
```{r}
sum_activity <- sum(is.na(activity_data))
```

```{r}
print(sum_activity)
```

### Histograms of NAs repartition per time interval and per date:
```{r}
missing_val <- subset(activity_data, is.na(steps))
par(mfrow = c(2,1), mar = c(2, 2, 1, 1))

hist(missing_val$interval, main="NAs Repartition per Interval", col="yellow")
hist(as.numeric(missing_val$date), main = "NAs Repartion per Date", col="yellow", breaks = 61)
```

### Histograms of total number of steps taken each day after missing values are imputed:
```{r}
steps_per_interval_mean <- tapply(activity_data$steps, activity_data$interval, mean, na.rm = TRUE)

activity_NAs <- activity_data[is.na(activity_data$steps),]
activity_non_NAs <- activity_data[!is.na(activity_data$steps),]
activity_NAs$steps <- as.factor(activity_NAs$interval)

levels(activity_NAs$steps) <- steps_per_interval_mean
levels(activity_NAs$steps) <- round(as.numeric(levels(activity_NAs$steps)))

activity_NAs$steps <- as.integer(as.vector(activity_NAs$steps))

imputed_activity <- rbind(activity_NAs, activity_non_NAs)

par(mfrow = c(1,2))

steps_day <- aggregate(steps ~ date, data = activity_data, FUN = sum, na.rm = TRUE)
hist(steps_day$steps, xlab = "Steps per Day", main = "NAs Removed - Total steps/day", col="yellow")

imp_activity_steps_day <- aggregate(steps ~ date, data = imputed_activity, FUN = sum, na.rm = TRUE)
hist(imp_activity_steps_day$steps, xlab = "Steps per Day", main = "NAs Imputed - Total steps/day", col="yellow")
```

##### Calculate and compare the mean and median values when NAs are removed and NAs are imputed :
```{r}
imp_mean_steps <- mean(imp_activity_steps_day$steps)
imp_median_steps <- median(imp_activity_steps_day$steps)

imp_mean_steps <- format(imp_mean_steps,digits=1)
imp_median_steps <- format(imp_median_steps,digits=1)

results_mean_median <- data.frame(c(mean_steps, median_steps), c(imp_mean_steps, imp_median_steps))

colnames(results_mean_median) <- c("NA removed", "Imputed NA")
rownames(results_mean_median) <- c("Mean", "Median")

xt <- xtable(results_mean_median)
```

```{r xtable, results="asis"}
print(xt, type="html")
```

### Panel plot comparing the average number of steps taken per 5-min. interval across weekdays and weekends:
```{r}
imputed_activity$dayType <- ifelse(weekdays(as.Date(imputed_activity$date)) == "Saturday" | weekdays(as.Date(imputed_activity$date)) == "Sunday", "weekend", "weekday")
imputed_activity$dayType <- factor(imputed_activity$dayType)

steps_interval_dayType <- aggregate(steps ~ interval + dayType, data = imputed_activity, FUN = mean)

head(steps_interval_dayType)

names(steps_interval_dayType) <- c("interval", "day_type", "mean_steps")

s <- ggplot(steps_interval_dayType, aes(interval, mean_steps,color = day_type)) + geom_line() + facet_grid(day_type~.) + labs(x = "Intervals", y = "Average Steps", title = "Activity")

print(s)

```














